---
title: "Project I"
author: "Yuri Freitas"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---

# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Dataset: ECOM
# Data understanding: Profiling -> Data quality
# Exploratory data analysis -> Variable kind, 
#                              Distributions (histogram, box plot), have a unique table (view)
#                              Basic KPI: Sells, 
#                                         quantity, 
#                                         count status
# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# Clean lists
```{r}
rm(list = ls())
graphics.off()
```

# Load necessary library
```{r}
library(dplyr)
library(ggplot2)
```

# Read from ECOM data source
```{r}
#LOADING  CUSTOMER
url_path <- paste0("https://docs.google.com/spreadsheets/d/e/2PACX-1vQIGXZHYrUMsj1e1cCi3JCZWHawxfaTuT3OX1_iTLQUt2tJKQ58jErOIhCGat9HSplMBMxAsFQxSBos/pub?output=csv")
customer <- read.csv(url_path, sep = ",")
rm(url_path)
#LOADING ORDER_ITEM
url_path <- paste0("https://docs.google.com/spreadsheets/d/e/2PACX-1vTpbThjdbYWHKFrJoqr0NkNQaaEneFl_1a8u1ki9PxmForQ0rnS7-4RXflKx0KHfz4fVA1DgrBLIpi9/pub?output=csv")
order_item <-read.csv(url_path, sep = ",")
rm(url_path)
#LOADING PRODUCTS
url_path <- paste0("https://docs.google.com/spreadsheets/d/e/2PACX-1vS-UgEDZPC9y4Z5SSZF-D0IfRGt_OIoq8EcuZF0t8qaIbErQBHf5zxDcJvwTo8kOzfx4csTqqiIN9fr/pub?output=csv")
product <- read.csv(url_path, sep = ",")
rm(url_path)
#LOADING PAYMENTS
url_path <-paste0("https://docs.google.com/spreadsheets/d/e/2PACX-1vRUd_uWktNGjPmFqxkpl1Wp8G1v2wVXFscFj8MqgaXQoBnqsZfDPVp6dmTBxdaSLLO4nL4-9ABhly6i/pub?output=csv")
payment <- read.csv(url_path, sep = ",")
rm(url_path)
#LOADING ORDER
url_path <-paste0("https://docs.google.com/spreadsheets/d/e/2PACX-1vSWKMH7XtzfEnHf9_8SZpiMBXD8z5VnUAvVF8WrN6KB-6FAOkYMoNQrRu9CKW4H_IHNTkDowKKxLMJu/pub?output=csv")
order <- read.csv(url_path, sep =",")
rm(url_path)
```

# Counts in the tables
```{r}
summary(order)
summary(order_item)
summary(customer)
summary(product)
summary(payment)
```

```{r}
# View(order)
# View(order_item)
# View(product)
# View(customer)
# View(payment)
```

# Getting only interesting fields for analysis
```{r}
order <- order %>% select(order_id, customer_id, order_status, order_purchase_timestamp, order_delivered_customer_date)
order_item <- order_item %>% select(order_id, order_item_id, product_id, price, freight_value) 
order_order_item <- inner_join(order, order_item, by = "order_id")
```

# Some keys to check the data anytime
```{r}
#order_id for check payments = '009ac365164f8e06f59d18a08045f6c4'
#cutomer_unique_id with many orders= '004288347e5e88a27ded2bb23747066c'
#order_id with many itens order_id=='001ab0a7578dd66cd4b0a71f5b6e1e41'
#first order to customer_unique_id = '2f64e403852e6893ae37485d5fcacdaf'
#last order to customer_unique_id = '24ac2b4327e25baf39f2119e4228976a'
#order with many products distincts = '8272b63d03f5f79c56e9e4120aec44ef'
```

# Create a data_customer df with all customers and its orders
```{r}
data_customer <- inner_join(order, customer, by = "customer_id") %>% select(customer_unique_id, customer_city, customer_state, order_id, order_status, order_purchase_timestamp, order_delivered_customer_date)
```

# Enriching the data_customer with number of itens in each order, total price, total freight value and price + freight
```{r}
data_customer <- data_customer %>% inner_join(select(order_order_item, order_id, price, freight_value), by = "order_id") %>% 
                  group_by(customer_unique_id, order_id) %>% mutate(order_itens = n(), total_price = sum(price), total_freight_value = sum(freight_value)) %>% 
                  select(-price, -freight_value) %>% 
                  distinct()
data_customer$total_price_freight <- data_customer$total_price + data_customer$total_freight_value
```

# Enriching the data with total payment from payment and also the number of payments, sometimes it has many payments for the same order (vouher, credit card) etc. 
```{r}
data_customer <- data_customer %>% 
                  inner_join(select(payment, order_id, payment_value), by = "order_id") %>% 
                  group_by(order_id) %>% 
                  mutate(qty_payments = n(), total_payment = sum(payment_value)) %>% 
                  select(-payment_value) %>% 
                  distinct()
```

# Cast chr to datetimestamp and get the weekday for purchase and delivered
```{r}
data_customer$order_purchase_timestamp <- as.POSIXct(data_customer$order_purchase_timestamp, format = "%Y-%m-%d %H:%M:%S")
data_customer$order_delivered_customer_date <- as.POSIXct(data_customer$order_delivered_customer_date, format = "%Y-%m-%d %H:%M:%S")
data_customer$purchase_weekday <- weekdays(data_customer$order_purchase_timestamp)
data_customer$delivered_weekday <- weekdays(data_customer$order_delivered_customer_date)
```

# Create a specific df to understanding products
```{r}
data_product <- data_customer %>% inner_join(select(order_order_item, order_id, product_id), by = "order_id") %>% 
                  inner_join(select(product, product_id, product_category_name), by = "product_id") %>%
                  select(order_id, product_id, product_category_name)
```

# Orders per weekday
```{r}
orders_per_weekday <- data_customer %>%
                      group_by(purchase_weekday) %>%
                      summarise(number_of_orders_weekday = n())

# Define a color palette for weekdays
weekday_colors <- c("Sunday" = "#FF9999",   # Light red
                    "Monday" = "#FFCC99",  # Light orange
                    "Tuesday" = "#FFFF99", # Light yellow
                    "Wednesday" = "#CCFF99", # Light green
                    "Thursday" = "#99CCFF",  # Light blue
                    "Friday" = "#CC99FF",    # Light purple
                    "Saturday" = "#FF99FF")  # Light pink

# Plot the number of orders per weekday with different colors
ggplot(data_customer, aes(x = purchase_weekday, fill = purchase_weekday)) +
  geom_bar() +
  scale_fill_manual(values = weekday_colors) +
  labs(title = "Number of Orders per Weekday",
       x = "Weekday",
       y = "Number of Orders") +
  theme_minimal() +
  theme(axis.text.x = element_text(face = "bold")) +
  geom_hline(yintercept = mean(orders_per_weekday$number_of_orders_weekday), linetype = "dashed", color = "black") +
  geom_text(aes(x = 4, y = mean(orders_per_weekday$number_of_orders_weekday)+400, label = "Mean"), color = "black", fontface = "bold")
```

# Orders per City
```{r}
orders_per_city <- data_customer %>%
                   group_by(customer_city) %>%
                   summarise(number_of_orders_city = n())

# Select only the top 7 cities
top_7_cities <- orders_per_city %>%
                arrange(desc(number_of_orders_city)) %>%
                head(7)

# Plot the number of orders per weekday with different colors
ggplot(top_7_cities, aes(x = reorder(customer_city, -number_of_orders_city), y = number_of_orders_city, fill = customer_city)) +
  geom_bar(stat = "identity") +
  labs(title = "Number of Orders per City (Top 7)",
       x = "City",
       y = "Number of Orders") +
  theme_minimal() +
  theme(axis.text.x = element_text(face = "bold"))
```

# We can see the number of orders in 2018 is decreasing generaly
```{r}
# Aggregate data by day
data_customer_agg <- data_customer %>%
  mutate(date = as.Date(order_purchase_timestamp)) %>%
  group_by(date) %>%
  summarise(orders = n())

# Plot the chart with trend line starting from 2018
ggplot(data = data_customer_agg, aes(x = date, y = orders)) +
  geom_line(color = "steelblue") +
  geom_point(color = "steelblue") +
  geom_smooth(data = filter(data_customer_agg, date >= as.Date("2018-01-01")),
              aes(x = date, y = orders),
              method = "lm", se = FALSE, color = "red") +  # Add trend line starting from 2018
  labs(title = "Number of Orders Over Time",
       x = "Date",
       y = "Number of Orders") +
  theme_minimal()
```

# Top 7 products by orders
```{r}
# Count the number of orders for each product category
category_counts <- table(data_product$product_category_name)

# Convert the result to a data frame and sort it by counts
category_counts_df <- as.data.frame(category_counts)
category_counts_df <- category_counts_df[order(-category_counts_df$Freq), ]

# Select top 7 categories
top_7_categories <- category_counts_df[1:7, ]

# Plot the bar chart for top 7 categories
ggplot(top_7_categories, aes(x = reorder(Var1, Freq), y = Freq)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(title = "Top 7 Product Categories by Sales",
       x = "Product Category",
       y = "Number of Orders") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
  coord_flip()
```

# Remove data with null values
```{r}
#data_customer <- na.omit(data_customer)
#View(data_customer)
```

#visualizing NA
```{r}
colSums(is.na(data_customer))
```

#Calculating the average delivery time
```{r}
 order_delivered_customer_date
data_customer$delivered_time <- difftime(data_customer$order_delivered_customer_date, data_customer$order_purchase_timestamp, units = "days")


mean_delivered_time <- mean(data_customer$delivered_time, na.rm = TRUE)


data_customer$mean_delivered_time <- mean_delivered_time
```

#Replace the values in the Order_Delivered_Customer_date column with calculated values
```{r}

data_customer$order_delivered_customer_date<- data_customer$order_purchase_timestamp + mean_delivered_time


data_customer$order_delivered_customer_date[is.na(data_customer$order_delivered_customer_date)] <- data_customer$order_delivered_customer_date[is.na(data_customer$order_delivered_customer_date)]

```

# Filling the values in the column Delivered_wekday with the average
```{r}
day_week <- c("segunda-feira", "terça-feira", "quarta-feira", "quinta-feira", "sexta-feira", "sábado", "domingo")

data_customer$purchase_weekday_num <- match(data_customer$purchase_weekday, day_week)
data_customer$delivered_weekday_num <- match(data_customer$delivered_weekday, day_week)

mean_weekday <- round(mean(c(data_customer$purchase_weekday_num, data_customer$delivered_weekday_num), na.rm = TRUE))


mean_weekday_str <- day_week[mean_weekday]

data_customer$delivered_weekday[is.na(data_customer$delivered_weekday)] <- mean_weekday_str
```

#Filling NA with mean from order_purchase_timestamp
#visualizing NA
```{r}

mean_buy <- mean(data_customer$order_purchase_timestamp, na.rm = TRUE)
mean_day_week <- weekdays(mean_buy)

data_customer$order_purchase_timestamp[is.na(data_customer$order_purchase_timestamp)] <- mean_buy


data_customer$purchase_weekday[is.na(data_customer$purchase_weekday)] <- mean_day_week


mean_delivered_time <- mean(data_customer$delivered_time, na.rm = TRUE)
data_customer$delivered_time[is.na(data_customer$delivered_time)] <- mean_delivered_time


data_customer <- subset(data_customer, select = -c(purchase_weekday_num, delivered_weekday_num))


colSums(is.na(data_customer))

```

