---
title: "Project I"
author: "Yuri Freitas"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---

# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Exploratory analysis
#   - dataset: ECOM
# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# Clean lists
```{r}
rm(list = ls())
graphics.off()
```

# Load necessary library
```{r}
library(dplyr)
```

# Read from ECOM data source
```{r}
order <- read.csv("C:/Users/yfreitas/Documents/RSamples/ProjectI/dataset_projeto1_20240323/fct_orders.csv")
customer <- read.csv("C:/Users/yfreitas/Documents/RSamples/ProjectI/dataset_projeto1_20240323/dim_customers.csv")
order_item <- read.csv("C:/Users/yfreitas/Documents/RSamples/ProjectI/dataset_projeto1_20240323/fct_order_items.csv")
product <- read.csv("C:/Users/yfreitas/Documents/RSamples/ProjectI/dataset_projeto1_20240323/Dim_products.csv")
payment <- read.csv("C:/Users/yfreitas/Documents/RSamples/ProjectI/dataset_projeto1_20240323/dim_payments.csv")
```

# Counts in the tables
```{r}
summary(order)
summary(order_item)
summary(customer)
summary(product)
summary(payment)
```

```{r}
# View(order)
# View(order_item)
# View(product)
# View(customer)
# View(payment)
```

# Getting only interesting fields for analysis
```{r}
order <- order %>% select(order_id, customer_id, order_status, order_purchase_timestamp, order_delivered_customer_date)
order_item <- order_item %>% select(order_id, order_item_id, product_id, price, freight_value) 
order_order_item <- inner_join(order, order_item, by = "order_id")
```

# Checking for 1 order in specific
```{r}
#order_id for check payments = '009ac365164f8e06f59d18a08045f6c4'
#cutomer_unique_id with many orders= '004288347e5e88a27ded2bb23747066c'
#order_id with many itens order_id=='001ab0a7578dd66cd4b0a71f5b6e1e41'))
```

# Create a data_customer df with all customers and its orders
```{r}
data_customer <- inner_join(order, customer, by = "customer_id") %>% select(customer_unique_id, order_id, order_status, order_purchase_timestamp, order_delivered_customer_date)
```

# Enriching the data_customer with number of itens in each order, total price, total freight value and price + freight
```{r}
data_customer <- data_customer %>% inner_join(select(order_order_item, order_id, price, freight_value), by = "order_id") %>% 
                  group_by(customer_unique_id, order_id) %>% mutate(order_itens = n(), total_price = sum(price), total_freight_value = sum(freight_value)) %>% 
                  select(-price, -freight_value) %>% 
                  distinct()
data_customer$total_price_freight <- data_customer$total_price + data_customer$total_freight_value
```

# Enriching the data with total payment from payment and also the number of payments, sometimes it has many payments for the same order (vouher, credit card) etc. 
```{r}
data_customer <- data_customer %>% 
                  inner_join(select(payment, order_id, payment_value), by = "order_id") %>% 
                  group_by(order_id) %>% 
                  mutate(qty_payments = n(), total_payment = sum(payment_value)) %>% 
                  select(-payment_value) %>% 
                  distinct()
```

```{r}
View(data_customer)
```

