# Customer Segmentation: RFM (Recency, Frequency, Monetary)
# Low-Value: Less active customers — customers who rarely make purchases and generate low revenues
# Mid-Value: Those who use the platform quite frequently and generate moderate revenues
# High-Value: Very active customers, who spend very often and spend more than any other customers

# Frequency
# We see that 90% of customers have only 1 order, so, in this case I am think if it worth to use k-means to create cluster
# For this case I think we could only spread in 2 cluster, the ones with 1 order and other with more orders
```{r}
counts <- data_customer_segmentation %>%
          group_by(count_orders) %>%
          summarise(count = n())

counts
```

# Simple distribution between cluster 1 and 2
```{r}
data_customer_segmentation <- data_customer_segmentation %>% 
                            mutate(frequency_cluster = ifelse(count_orders == 1, 1, 2))
```

```{r}
counts <- data_customer_segmentation %>%
          group_by(frequency_cluster) %>%
          summarise(count = n())

counts
```