
# Customer Segmentation: RFM (Recency, Frequency, Monetary)
# Recency: Number of days from the last purchase untill 01-01-2018 (time without buy)
# Frequency: Number of orders
# Monetary: Amount spent
# Low-Value: Less active customers — customers who rarely make purchases and generate low revenues
# Mid-Value: Those who use the platform quite frequently and generate moderate revenues
# High-Value: Very active customers, who spend very often and spend more than any other customers


# Recency
# Calculate how many days it has until 01-01-2018
```{r}
data_customer_segmentation <- data_customer_cutoff %>%
                              group_by(customer_unique_id, target) %>%
                              summarise(count_orders = n(), max_purchase = max(order_purchase_timestamp), total_paid = sum(total_payment)) %>%
                              arrange(max_purchase) %>%
                              select(customer_unique_id, target, total_paid, count_orders, max_purchase)
data_customer_segmentation$recency <- as.numeric(difftime(as.Date("2018-01-01"), data_customer_segmentation$max_purchase, units = "days"))
```

# Distribution of the recency
```{r}
ggplot(data_customer_segmentation, aes(x = recency)) +
  geom_histogram(binwidth = 10, fill = "skyblue", color = "black") +
  labs(title = "Recency Histogram",
       x = "Recency (Days)",
       y = "Count")
```

# To assign a recency score, we’ll be using KMeans clustering. But before that, as with other KMeans projects, we need to know how many clusters we need.
# Also, we'll use only use numeric features
```{r}
customer_segmentation_recency_num <- data_customer_segmentation[,6]
head(customer_segmentation_recency_num)
```

# As K-means calculates distances, we'll standardize the numerical variables
```{r}
customer_segmentation_recency_std <- as.data.frame(scale(customer_segmentation_recency_num))
head(customer_segmentation_recency_std)
```

# Note that means=0 and sd=1 for all variables after standardization
```{r}
summary(customer_segmentation_recency_std)
sapply(customer_segmentation_recency_std, sd)
```

# Determine Number of Clusters --------------------------------------------
```{r}
# Get just a sample of data, these algorithms are heavy
sampled_data_recency_std <- customer_segmentation_recency_std %>%
                            sample_n(10000, replace = FALSE)

# "Elbow method", for total within-cluster sum of square
factoextra::fviz_nbclust(sampled_data_recency_std, kmeans, method = "wss")

# "Silhouette-score method", for how similar a data point is within-cluster
factoextra::fviz_nbclust(sampled_data_recency_std, kmeans, method = "silhouette")
```

# Define number of clusters
```{r}
k_clusters <- 4
```

# Apply K-means algorithm
```{r}
# Perform k-means clustering
set.seed(811)
fit.kmeans <- kmeans(customer_segmentation_recency_std,
                     centers = k_clusters,
                     nstart = 20, # how many random sets should be chosen
                     iter.max = 100)
print(fit.kmeans)
```

# Note that "Cluster means" are the centers of each cluster ->
```{r} 
fit.kmeans$centers
```

# Assigne the cluster number to data_customer_segmentation and show the numbers grouped by recency_cluster
```{r}
# Create a mapping from the original cluster centers to the ordered centers
cluster_mapping <- match(fit.kmeans$centers, rev(sort(fit.kmeans$centers)))
# Append cluster labels to the standardized data
customer_segmentation_recency_std$cluster <- cluster_mapping[fit.kmeans$cluster]
# Append cluster and recency to data_customer_segmentation
data_customer_segmentation$recency_cluster <- customer_segmentation_recency_std$cluster
data_customer_segmentation$recency_std <- customer_segmentation_recency_std$recency

counts <- data_customer_segmentation %>%
          group_by(recency_cluster) %>%
          summarise(count = n())

counts
```

# Plot observations assigned to centers
```{r}
plot(recency ~ 1,
     data = data_customer_segmentation,
     col = fit.kmeans$cluster + 1,
     pch = 19,
     main = "Number of orders by Recency (with assigned clusters)",
     sub = "Note: These are 4 clusters, classification for recency variable",
     xlab = "Number of orders",
     ylab = "Recency")
```

# Plot a bar chart showing the cluster numbers
```{r}
ggplot(data_customer_segmentation, aes(x = recency_cluster)) +
  geom_bar(fill = "skyblue", color = "black") +
  labs(title = "Counts of Recency Clusters",
       x = "Recency Cluster",
       y = "Count")
```
